---
title: "Data605wk12"
author: "Charls Joseph"
date: "April 21, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RCurl)
library(knitr)
options(scipen=5)
```

## R Markdown

```{r}

who_data <- read.csv(text = getURL("https://raw.githubusercontent.com/charlsjoseph/data605wk12/master/who.csv"), header = T, stringsAsFactors =F)

kable(who_data[sample(nrow(who_data), 10), ], align='l', row.names=FALSE)

```

```{r}
summary(who_data)
```

### Qn1.  Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.

```{r}
who.lm <- lm(LifeExp ~ TotExp, data = who_data)
plot(who_data$TotExp, who_data$LifeExp, xlab = 'Total Expenditure - $', ylab = 'Average Life Expectancy- in Years', main='Avg. Life Expectancy Vs. Expenditure')
abline(who.lm, col="red")

```

### The scatter plot doesnt fit properly into the linear line. It shows the expenditure is constant until the age 70-75 and slowly increases beyond 70. 

Lets find the summary of linear regression and analyze the key parameters. 


```{r}

summary(who.lm)
```
Regression model 

$$ Avg Life Expectancy=64.75337453+0.00006297∗Total Expenditure $$
The model explains on every unit of increase in Total expenditure, the life expenditure goes up by 0.00006297.

```{r}

plot(who.lm, which=c(1,1))

```

The resduals are not scattererd, but having an unsual pattern which tells that relation doesnt hold the linear behaviour. 
P-value is very less which explains there is a strong relation between LifeExp and TotExp.
Adjusted R^2 and R^2 is very less which explains the totalExp is not only an explanatory variable and model can be possibly more accurate or better. 


### 2. Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is “better?”

```{r}
#raise total expenditure to the power of 0.06
who_data$TotExp_1 <- (who_data$TotExp)^0.06

#raise life expectancy to the power of 4.6
who_data$LifeExp_1 <- (who_data$LifeExp)^4.6

#build new linear model
who_1.lm <- lm(LifeExp_1 ~ TotExp_1, data = who_data)
plot(who_data$TotExp_1, who_data$LifeExp_1, xlab = 'Total Expenditure(in US Dollars) raised to 0.06', ylab = 'Average Life Expectancy(in Years) raised to 4.6', main='Avg. Life Expectancy Vs. Expenditure')
abline(who_1.lm, col="red")
      
```

This explains a better linear relation. Lets check the model key paramaters. 
```{r}
summary(who_1.lm)
```
Regression model is 

$$LifeExp1=−736527909+620060216∗TotExp1 $$



```{r}
plot(who_1.lm, which=c(1,1))

```


The residual looks to be scatter and model looks to follow the linear pattern. 
Adjusted R^2 and R^2 looks to be closer to 1 and better than previous model. 

### 3. Using the results from 2, forecast life expectancy when TotExp^.06 =1.5. Then forecast life expectancy when TotExp^.06=2.5.

```{r}
result1 <- (-736527909 + (620060216 * 1.5))^(1/4.6)
result2 <- (-736527909 + (620060216 * 2.5))^(1/4.6)
result1
result2
```

Estimated life expectancy is 63.31153 years when TotExp(0.06) is 1.5.
Estimated life expectancy is 86.50645 years when TotExp(0.06) is 2.5.

### 4. Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model? LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp


```{r}
who_2.lm <- lm(LifeExp ~ PropMD + TotExp + PropMD * TotExp, data = who_data)
summary(who_2.lm)

```
Regression model is 
$$
Average Life Expectancy=62.77270326+1497.49395252×PropMD+0.00007233×TotExp−0.00602569×PropMD∗TotExp $$



p-Value for PropMD, TotExp and PropMD * TotExp are highly relevant to estimate average life expectancy. 
R2 and Adj. R2 values decreased compared to earlier 

###5. Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?

```{r}
TotExp = 14
PropMD = 0.03
forcast.life.exp = 62.77270326 + (1497.49395252 * PropMD) + (0.00007233 * TotExp) - (0.00602569 * PropMD * TotExp)
forcast.life.exp


```


The result doesnt looks to be realistic. This could be due to in-efficiency of the model variables.


